---
title: "Simulations OVB/type M trade-off: IV"
author:
  - name: Vincent Bagilet 
    url: https://vincentbagilet.github.io/
    affiliation: Columbia University
    affiliation_url: https://www.columbia.edu/
date: "`r Sys.Date()`"
output: 
  distill::distill_article:
    toc: true
editor_options: 
  chunk_output_type: console
---

<style>
body {
text-align: justify}
</style>

```{r setup, include=FALSE, results='hide', warning=FALSE}
library(knitr)
opts_chunk$set(fig.path = "images/",
               cache.path = "cache/",
               cache = FALSE,
               echo = TRUE, #set to false to hide code
               message = FALSE,
               warning = FALSE,
               out.width = "85%",
               dpi = 200,
               fig.align = "center")  
```  

## Purpose of the document

In this document, I run a simulation exercise to illustrate issues arising when estimating Two-Way Fixed Effect (TWFE) models with staggered and heterogeneous treatment.

First, I load useful packages. Note that some packages are optional. If you do not want to install/use them, you may need to modify part of your code. `mediocrethemes` is my ggplot theme package. If you want to use it, you can find instructions [here](https://vincentbagilet.github.io/mediocrethemes/).

```{r packages, message=FALSE, warning=FALSE}
#necessary packages
library(tidyverse) 
library(fixest)
library(knitr) 
library(broom)

#Optional packages
library(mediocrethemes) 
library(tictoc) 
library(here)
library(beepr)

set.seed(1)

mediocrethemes::set_mediocre_all(pal = "coty")
```

## Building the example

To illustrate issues arising with TWFE, I build simple fake data, estimate a TWFE model on it and replicate the analysis several times.

### Modelisation choices

To simplify, I consider the assumptions described below. Of course these assumptions are purely arbitrary and I invite you to play with them. Note that, fixed effects and the covariate are not necessary to the analysis. I only add them to make the analysis more realistic if necessary but I set their baseline values to 0.

- Each individual has fixed characteristics drawn from a normal distribution,
- Each time period presents fixed characteristics also drawn from a normal distribution,
- A unique covariate $x$ drawn from a normal distribution,
- A proportion $p_{treat}$ of individuals are ever treated over the period. Hence, a proportion of $1-p_{treat}$ individuals are never treated over the period. I draw these individual at random. Note that the value of the individual identifiers do not matter here. So I could assume that the non-treated individuals are those with the larger individual ids for instance,
- The implementation of the treatment can be staggered or not. If it is not staggered, the treatment date is set to be in the middle of the period
- The treatment can vary along two dimensions, time and individual. Details are given below.

More precisely, I set: 

- $N_i$ the number of individual
- $N_t$ the number of periods
- $\lambda_i \sim \mathcal{N}(\mu_{IFE}, \sigma_{IFE}^{2})$ the fixed effect for individual $i$
- $\eta_t \sim \mathcal{N}(\mu_{TFE}, \sigma_{TFE}^{2})$ the fixed effect for time period $t$
- $x_{it} \sim \mathcal{N}(\mu_{x}, \sigma_{x}^{2})$
- $e_{it} \sim \mathcal{N}(0, \sigma_{e}^{2})$ some noise
- $T_{it}$ represent the treatment allocation, it is equal to one if individual $i$ is treated at time $t$ and 0 otherwise,
- $y_{it} = \alpha + \beta_i T_{it} + \gamma x_{it} + \lambda_i + \eta_t + e_{it}$ where $\alpha$ and $\gamma$ are some constants.
- $\beta_i$ is represents the magnitude of the treatment effect and is linked to the input parameter `beta`. Across individuals, the treatment can either be homogeneous (`het_indiv == homogeneous`, for each individual, the treatment is equal to `beta`), random (`het_indiv == random`, for each individual, the treatment is drawn from  $\mathcal{U}(0.5\beta, 1.5\beta)$) or larger for those that are treated first (`het_indiv == large_first`, for each individual, the treatment is equal to $N_t - \beta$). Across time, the effect of the treatment can either be constant (`het_time == constant`) or increasing linearly in time (`het_time == linear`).

I also create a bunch of variables that can be useful:
- $InTreatment_i$ equal to 1 if individual $i$ ever gets treated,
- $t^{event}_i$ equal to the date at which individual $i$ gets treated,
- $t^{centered}_i$ representing the distance in terms of period to the beginning of the treatment for individual $i$,
- $Post_{it}$ equal to 1 if the period $t$ is after the treatment has begun for individual $i$. This variable is only useful for non-staggered treatment allocation,

### Data generation

I write a simple function that generates the data. It takes as input the values of the different parameters and returns a data frame containing all the variables for this analysis. 

```{r DGP}
generate_data_TWFE <- function(N_i,
                               N_t,
                               sigma_e,
                               p_treat,
                               staggered,
                               het_indiv,
                               het_time,
                               alpha,
                               beta,
                               mu_indiv_fe = 0, 
                               sigma_indiv_fe = 0,
                               mu_time_fe = 0, 
                               sigma_time_fe = 0,
                               mu_x = 0, 
                               sigma_x = 0,
                               gamma = 0
                             ) {

  if (!is.logical(staggered)) {stop("staggered must be logical")} 
  if (!(het_indiv %in% c("large_first", "random", "homogeneous"))) {
    stop('het_indiv must be either "large_first", "random" or "homogeneous"')
  } 
  if (!(het_time %in% c("constant", "linear"))) {
    stop('het_time must be either "constant" or "linear"')
  } 
  
  data <- tibble(indiv = 1:N_i) %>%
    mutate(in_treatment = (indiv %in% sample(1:N_i, floor(N_i*p_treat)))) %>% 
    crossing(t = 1:N_t) %>%
    group_by(indiv) %>%
    mutate(
      indiv_fe = rnorm(1, mu_indiv_fe, sigma_indiv_fe),
      t_event = ifelse(staggered, sample(2:(N_t - 1), 1), floor(N_t/2)), 
        #I use 2:(N_t-1) to have a pre and post period
      t_event = ifelse(in_treatment, t_event, NA),
      beta_i = case_when(
        het_indiv == "large_first" ~ N_t-t_event,
        het_indiv == "random" ~ runif(1, beta*0.5, beta*1.5), 
        het_indiv == "homogeneous" ~ beta
      ),
      beta_i = ifelse(is.na(t_event), 0, beta_i)
    ) %>%
    ungroup() %>%
    group_by(t) %>%
    mutate(time_fe = rnorm(1, mu_time_fe, sigma_time_fe)) %>%
    ungroup() %>%
    mutate(
      post = (t > t_event),
      treated = in_treatment & post, 
      beta_i = ifelse(
        het_time == "linear" & post & !is.na(t_event),
        beta_i*(t - t_event), 
        beta_i
      ),
      t_centered = t - t_event,
      x = rnorm(nrow(.), mu_x, sigma_x),
      e = rnorm(nrow(.), 0, sigma_e),
      y0 = alpha + gamma * x + indiv_fe + time_fe + e,
      y1 = y0 + beta_i,
      y = treated*y1 + (1 - treated)*y0
    )
  
  return(data)
}
```

I set baseline values for the parameters as very standard. These values are completely arbitrary.

```{r baseline_param}
baseline_parameters_TWFE <- tibble(
  N_i = 20,
  N_t = 50,
  sigma_e = 1,
  p_treat = 0.8,
  staggered = TRUE,
  het_indiv = "homogeneous",
  het_time = "constant",
  alpha = 1,
  beta = 1
)
```

Here is an example of data created with the data generating process and baseline parameter values, for 2 individuals and 8 time periods:

```{r example_data, echo=FALSE}
baseline_parameters_TWFE %>% 
  mutate(N_i = 2, N_t = 8) %>%
  pmap_dfr(generate_data_TWFE) %>% #use pmap to pass the set of parameters
  select(indiv, t, y, in_treatment, post, treated, t_centered, e) %>% 
  kable()
```

Let's now have a look at different types of treatment and treatment allocations. First, let's look at treatment allocation mechanisms. The allocation can either be staggered or not

```{r treatment_allocation, echo=FALSE}
labs_graph_staggered <- labs(
    title = "Treatement assignment across time and individuals",
    x = "Time index", 
    y = "Individual id", 
    fill = "Treated"
  )

baseline_parameters_TWFE %>% 
  mutate(staggered = FALSE) %>%
  pmap_dfr(generate_data_TWFE) %>% #use pmap to pass the set of parameters
  ggplot(aes(x = t, y = factor(indiv), fill = fct_rev(factor(treated)))) + 
  geom_tile(color = "white", lwd = 0.5, linetype = 1) +
  coord_fixed() +
  labs_graph_staggered + 
  labs(subtitle = "Non staggered")

baseline_parameters_TWFE %>% 
  mutate(staggered = TRUE) %>%
  pmap_dfr(generate_data_TWFE) %>% #use pmap to pass the set of parameters
  ggplot(aes(x = t, y = factor(indiv), fill = fct_rev(factor(treated)))) + 
  geom_tile(color = "white", lwd = 0.5, linetype = 1) +
  coord_fixed() +
  labs_graph_staggered + 
  labs(subtitle = "Staggered")
```

Now, let's vary treatment effect size across individuals, considering a staggered adoption.

```{r treatment_effect, echo=FALSE}
labs_graph_size <- labs(
    title = "Treatement effect size across time and individuals",
    x = "Time index", 
    y = "Individual id", 
    fill = "Treatment effect size"
  )

set_mediocre_all(pal = "blackandwhite")

baseline_parameters_TWFE %>% 
  mutate(het_indiv = "homogeneous", het_time = "constant") %>%
  pmap_dfr(generate_data_TWFE) %>% #use pmap to pass the set of parameters
  ggplot(aes(x = t, y = factor(indiv), fill = round(treated*beta_i, 2))) + 
  geom_tile(color = "white", lwd = 0.5, linetype = 1) +
  coord_fixed() +
  labs_graph_size + 
  labs(subtitle = "Homogeneous treatment effect across individuals, constant in time")

baseline_parameters_TWFE %>% 
  mutate(het_indiv = "random", het_time = "constant") %>%
  pmap_dfr(generate_data_TWFE) %>% #use pmap to pass the set of parameters
  ggplot(aes(x = t, y = factor(indiv), fill = round(treated*beta_i, 2))) + 
  geom_tile(color = "white", lwd = 0.5, linetype = 1) +
  coord_fixed() +
  labs_graph_size + 
  labs(subtitle = "Random treatment effect size across individuals, constant in time")

baseline_parameters_TWFE %>% 
  mutate(het_indiv = "large_first", het_time = "constant") %>%
  pmap_dfr(generate_data_TWFE) %>% #use pmap to pass the set of parameters
  ggplot(aes(x = t, y = factor(indiv), fill = round(treated*beta_i, 2))) + 
  geom_tile(color = "white", lwd = 0.5, linetype = 1) +
  coord_fixed() +
  labs_graph_size + 
  labs(subtitle = "First treated have larger treatment effect, constant in time")
```

A last thing we can vary is that we can make individual effects increase linearly in time.

```{r treatment_time, echo=FALSE}
baseline_parameters_TWFE %>% 
  mutate(het_indiv = "homogeneous", het_time = "linear") %>%
  pmap_dfr(generate_data_TWFE) %>% #use pmap to pass the set of parameters
  ggplot(aes(x = t, y = factor(indiv), fill = round(treated*beta_i, 2))) + #treated*beta_i
  geom_tile(color = "white", lwd = 0.5, linetype = 1) +
  coord_fixed() +
  mediocrethemes::scale_mediocre_c("blackandwhite") +
  labs_graph_size +
  labs(subtitle = "Treatment effect increasing linearly in time")

set_mediocre_all(pal = "coty")
```

### Estimation

After generating the data, we can run an estimation.

```{r estimate}
estimate_TWFE <- function(data) {
  reg <- data %>% 
    mutate(
      indiv = as.factor(indiv),
      t = as.factor(t),
      treated = as.numeric(treated),
      in_treatment = as.numeric(in_treatment),
      t_centered = as.factor(t_centered)
    ) %>% 
    feols(
    data = ., 
    fml = y ~ in_treatment:t_centered | indiv + t
    ) %>% 
    broom::tidy() %>% 
    rename(p_value = p.value, se = std.error) %>% 
    mutate(term = as.numeric(str_remove_all(term, "in_treatment\\:t_centered"))) %>% 
    rename(lag = term)
  
  return(reg)
}
```

```{r one_estimation}
baseline_parameters_TWFE %>% 
  pmap_dfr(generate_data_TWFE) %>%
  estimate_TWFE() 
```

### Run a whole simulation

 To do so I create the function `compute_simulation_TWFE`. This simple function takes as input the various parameters an returns a table with the estimate of the treatment, its p-value and standard error and all input parameters. 
 
But first, we want to compute the true effect (the ATT) to add it to our output.

```{r compute_true_effect}
compute_true_effect_TWFE <- function(data) {
  data %>% 
    filter(in_treatment) %>% 
    group_by(t_centered) %>% 
    summarise(true_effect = mean(treated*(y1 - y0))) %>% 
    rename(lag = t_centered)
}  
```

We can then compute the simulation.

```{r compute_simulation}
compute_simulation_TWFE <- function(N_i,
                                    N_t,
                                    sigma_e,
                                    p_treat,
                                    staggered,
                                    het_indiv,
                                    het_time,
                                    alpha,
                                    beta,
                                    mu_indiv_fe = 0,
                                    sigma_indiv_fe = 0,
                                    mu_time_fe = 0,
                                    sigma_time_fe = 0,
                                    mu_x = 0,
                                    sigma_x = 0,
                                    gamma = 0) {
  data <- generate_data_TWFE(
    N_i = N_i,
    N_t = N_t,
    sigma_e = sigma_e,
    p_treat = p_treat,
    staggered = staggered,
    het_indiv = het_indiv,
    het_time = het_time,
    alpha = alpha,
    beta = beta,
    mu_indiv_fe = mu_indiv_fe,
    sigma_indiv_fe = sigma_indiv_fe,
    mu_time_fe = mu_time_fe,
    sigma_time_fe = sigma_time_fe,
    mu_x = mu_x,
    sigma_x = sigma_x,
    gamma = gamma
  ) 
  
  data %>%
    estimate_TWFE() %>%
    mutate(
      N_i = N_i,
      N_t = N_t,
      sigma_e = sigma_e,
      p_treat = p_treat,
      staggered = staggered,
      het_indiv = het_indiv,
      het_time = het_time,
      alpha = alpha,
      beta = beta,
      mu_indiv_fe = mu_indiv_fe,
      sigma_indiv_fe = sigma_indiv_fe,
      mu_time_fe = mu_time_fe,
      sigma_time_fe = sigma_time_fe,
      mu_x = mu_x,
      sigma_x = sigma_x,
      gamma = gamma
    ) %>% 
    left_join(compute_true_effect_TWFE(data), by = "lag")
} 
```

Here is the output of one simulation: 

```{r one_simulation}
baseline_parameters_TWFE %>% 
  pmap_dfr(compute_simulation_TWFE) 
```

## Analysis of the results

To analyze the results, I build a simple function to run the regression and graph the results. It takes as inputs the baseline parameters and our parameters of interest.

```{r graph_result}
graph_results <- function(baseline_parameters, 
                          staggered = TRUE, 
                          het_indiv = "homogeneous", 
                          het_time = "constant") {
  
  baseline_parameters["staggered"] <- staggered
  baseline_parameters["het_indiv"] <- het_indiv
  baseline_parameters["het_time"] <- het_time
  
  graph <- baseline_parameters %>%
    pmap_dfr(compute_simulation_TWFE) %>% 
    filter(dplyr::between(lag, -5, 5)) %>% 
    mutate(
      estimate_level = (estimate - estimate[which(lag == 0)]),
      true_effect_level = (true_effect - true_effect[which(lag == 0)])
    ) %>% 
    ggplot() +
    geom_point(aes(x = lag, y = estimate_level)) +
    geom_point(aes(x = lag, y = true_effect_level), shape = 1) +
    # geom_errorbar(aes(x = lag, ymin=estimate_level-se, ymax=estimate_level+se)) +
    labs(
      x = "Lag",
      y = "Estimate (centered)",
      title = "Representation of estimates for each lag",
      subtitle = paste(
      ifelse(staggered, "Staggered,", "Non staggered,"),
      het_indiv, ",",
      het_time,
      "treatment"),
      caption = "Hollow points represent the centered true effect"
    )
  
  return(graph)
}
 
baseline_parameters_TWFE %>% 
  graph_results()
```

We can reproduce this graph for varying types of treatment.

```{r graph_result_others}
baseline_parameters_TWFE %>% 
  graph_results(staggered = TRUE, het_indiv = "large_first", het_time = "linear") 
```